Exploring a Red Wine Data Set by Julio Uribe
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(RColorBrewer)
library(gridExtra)
library(grid)
library(GGally)
library(scales)
library(memisc)
library(lattice)
library(MASS)
library(car)
library(reshape)
library(plyr)
library(dplyr)
library(wesanderson)
library(RCurl)
library(bitops)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the Data
getwd()
setwd("C:/Users/Jules/Desktop/Nano/EDA Project")
reds <- read.csv('wineQualityReds.csv', sep = ',')
```

### Introduction

This report explores a data set of red wine containing approximately 1600 
different kinds of wine. I'm a fan of red wine but wouldn't consider myself very knowledgeable. The purpose of this exploration is to learn more about what
chemical qualities influence overall quality of wine, rated by experts. Perhaps
we can take our conclusions to the real world and do some in field testing. 
Before exploring the data, I think that residual sugar, alcohol level, and 
acidity will have the most influence on quality of wine. overall I want to learn 
what affects the quality of wine so that I can make more informed predictions at 
what kind of wine I or others would enjoy

# Univariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
# first thing I wan to look at is what variables I have available and how the
# data set is organized
names(reds)
summary(reds)
head(reds, 10)
```

Our data set contains 12 variables plus an identifier varible with almost 1600 observations. These variables are not easily observed in a kitchen. For most of
these variables you'd need a chemistry set to extract the values. Let's remove
the x variable to focus on the other 12 variables.

```{r echo=FALSE, message=FALSE, warning=FALSE}
reds$X <- NULL

# Summary of Quality to get an idea of how
summary(reds$quality)

# histogram of quality to learn more about the spread
ggplot(reds, aes(x=factor(quality), y = ..count../sum(..count..))) + 
  geom_bar(fill = "firebrick3") +
  xlab('Wine Quality') +
  ylab('Wine Quality As A Proportion')
```

Wine quality distribution falls between a low of 3 and high of 8. We have a mean 
of 5.64 and a median of 6.00. Our max is 8 and min is 3. Most wine is rated 
either as a 5 or 6. When looking at the histogram of wine quality as a 
proportion we see that a little more than 40% of wine is rated a 5, about 40% is 
rated a 6, about 10% is rated a 7, and the rest of the values scatter out to 3, 
4, and 8.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Summary of Alcohol percentage
summary(reds$alcohol)

# Histogram for aclohol to see spread
ggplot(aes(alcohol), data = reds) +
  geom_histogram(fill = 'firebrick3') +
  xlab('Alcohol Content') +
  ylab('Count')
```

What does alcohol content look like when plotted? We see a positive skew with
most of our wine having about 9 percent and the tail decreases towards the 13 
percent mark with top percent of wines hitting almost 14 percent alcohol 
content. We have an average of 10.42 percent, a max of 14.90 percent, 
and minimum of 9.50. Our median is 10.20. We see a small peak below the 10 
percent mark with most of the distribution falling between about 9 percent up to
13 percent.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Summary for residual sugar
summary(reds$residual.sugar)

# Histogram for residual sugar
ggplot(aes(residual.sugar), data = reds) +
  geom_histogram(fill = 'firebrick3') +
  xlab('Residual Sugar') +
  ylab('Count')

ggplot(aes(residual.sugar), data = reds) +
  geom_histogram(fill = 'firebrick3') +
  scale_x_continuous(limits = c(0, 10), breaks = seq(0, 10, 2))+
  xlab('Residual Sugar') +
  ylab('Count')
# used limits here because we had a few outliers of relatively sweet wines >12.
```

When I first plotted residual sugar, we had a few outliers of relatively sweet
wines that came in at over 12g of sugar making this plot a bit difficult to see
so I used some limits to get a better view of our distribution. Here we see a
positive skew once again with the majority of our points falling between 1 to 3
grams of sugar. We have a mean of 2.539 and a median of 2.200.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Couple more univariate plots 
# Histogram for total sulfur dioxide
ggplot(aes(total.sulfur.dioxide), data = reds) +
  geom_histogram(fill = 'firebrick3') +
  xlab('Total Sulfure Dioxide') +
  ylab('Count')

# Histogram for volatile acidity
ggplot(aes(volatile.acidity), data = reds) +
  geom_histogram(fill = 'firebrick3') +
  xlab('Volatile Acidity') +
  ylab('Count')

# Density Histogram
ggplot(aes(density), data = reds) +
  geom_histogram(fill = 'firebrick3') +
  xlab('Density') +
  ylab('Count')
```

Plotting a few more variables to get a feel for the other variables. Total 
sulfur dioxide and volatile acidity both have positive skews as well. Our 
density histogram has a nice normal distribution centered at about 0.997. I'm
not exactly sure what affects the density but most wine tends to hang out at
about the same spot as water. Not too familiar with what causes changes in 
density but this difference in density probably won't give us much insight at 
surface level.


# Univariate Analysis

### What is the structure of your dataset?
The data set is structured in csv format. We have 1599 objects and 12 variables
not counting the "x" variable which is used to id the various wines. All of our 
variables are quantitative. We don't have any qualitative variables. Our quality
distribution is only 5 values(3, 4, 5, 6, 7, 8) which set us up for some easy
pairing to look at low quality, middle quality, and high quality wine.

### What is/are the main feature(s) of interest in your dataset?
We have various measurements for acid, sulfur, and ph scale related measures. 
I'll be focusing most of my analysis on the quality variable and how the other
variables affect this. Alcohol percentage is interesting. Does stronger alcohol
content mean worse taste? My gut feeling tells me stronger alcohol presence will
lead to poor taste and as result lower quality.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

My main goal is to learn what affects the overall quality of the wine. I'm 
gonna dig to see how the variables available to us are related to quality. I 
mentioned in the introduction that I think alcohol level, residual sugar, and 
acidity might be significant.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

I didn't have to manipulate the data too much since we're working with a clean
data set. I did remove the x variable since we don't have info on what the id's
correspond to. When creating a histogram for the residual sugar distribution, I 
used some limits to zoom in on the majority of the distribution. Otherwise it 
would have been tough to see ranges of our distribution.


# Bivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}
#ggcorr to scope out any strong correlations
ggcorr(reds)
```

We plot ggcorr to get an idea of our data and we see mostly weak correlations 
but there are a few low to medium sized correlations. Nothing as strong as we 
saw with the diamond data set in previous exercises. The strongest correlations 
I see in relation to wine quality are alcohol, volatile acidity, and sulphates. 
Other than wine quality, ther are high correlations for citric acid and fixed
acidity and volatile acidity.

```{r echo=FALSE, message=FALSE, warning=FALSE}
pair <- data.frame(
        Alcohol.Content = reds$alcohol,
        Volatile.Acidity = reds$volatile.acidity,
        Sulphates = reds$sulphates,
        Quality = reds$quality
)
ggpairs(pair, 
        lower = list(continous = wrap("points", shape = I('.'))),
        upper = list(combo = wrap("box", outlier.shape = I('.')))) +
  theme(
        axis.ticks = element_blank(),
        axis.text = element_blank())
cor.test(reds$citric.acid, reds$fixed.acidity)
cor.test(reds$citric.acid, reds$volatile.acidity)
```

In relation to quality, we get r= 0.476 for alcohol, r = -0.391 for volatile 
acidity, and r = 0.251 for sulphates. Alcohol correlation is fairly large in 
effect size so we'll start investigating that one first. Volatile acidity is 
moderate and sulphates is more on the small side. We also observed high 
correlations in the previous ggcorr plot for citric acid in relation to fixed 
acidity and volatile acidty at 0.67 and -0.55, respectively.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# alcohol and quality
ggplot(aes(factor(quality), alcohol), data = reds) +
  geom_boxplot(color = 'firebrick3') +
  xlab('Quality') +
  ylab('Alcohol Content')
```

It's interesting to see quality of wine positively correlated with alcohol. At
0.476 we have a decent correlation. Could higher alcohol content mean better
taste or is it a reflection of better brewing methods? Our boxplot here gives us 
an idea where the averages fall for each quantity value. However, it's difficult 
to see the correlation we observed with our initial ggcorr plot. There's a nice
upwards trend of the average alcohol content for quality values starting at 5 to
8. Our first two values of 3 and 4 don't appear to go with the trend very well.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# quality and alcohol and mean alcohol across quality
# plotting quality with alcohol means and Q1 and Q3
# also plotting geom smooth line to see gradual incrase in alcoholMean content
alc1 <- ggplot(aes(x = quality, y = alcohol), data = reds) +
  geom_point(alpha = 1/4, position = position_jitter(h = 0), color = 'firebrick3') +
  geom_line(stat = 'summary', fun.y = mean, color = 'black') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .25), linetype = 2, color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .75), linetype = 2, color = 'blue') +
  xlab('Quality') +
  ylab('Alcohol Content') 
alc1

alc1 <- alc1 + 
  scale_x_continuous(limits = c(3, 8), breaks = seq(0, 8, 1)) +
  scale_y_continuous(limits = c(0, 16), breaks = seq(0, 16, 2))
alc1
```

Here we made a scatter plot of quality and alcohol content but we also included
some averages to see if we can start seeing a correlation. Adding jitter and 
alpha better demonstrates the spread near the 5 and 6 values. We see the mean 
alcohol content hang around 10% at quality 5 and gradually increase to 12% as 
wine quality increases to 8. There doesn't seem to be much of a trend with 
quality ratings below 5. 

Our initial plot exaggerates the correlation because our y value for acohol 
content doesn't start at 0 so I adjusted to limits to better demonstrate what is 
going on. We see a bit of a spike and increase at quality rating 6 and higher. 
However, the correlation of 0.476 looks fairly weak on our plot. Perhaps if we 
grouped the data into low, medium, and high quality wine we would see a clearer 
relationship. For now I'll group the data by quality rating and get a single 
average for alcohol content at that rating.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# compute averages and store them into a new small data set.
redsByQuality <- reds %>%
  group_by(quality) %>%
  summarise(alcoholMean = mean(alcohol),
            alcoholMedian = median(alcohol),
            n = n()) %>%
  arrange(quality)

# plot the means in a smooth line
alc2 <- ggplot(aes(x = quality, y = alcoholMean), data = redsByQuality) +
  geom_point(color = 'firebrick3') +
  geom_smooth() +
  xlab('Quality') +
  ylab('Alcohol Content Mean') +
  scale_x_continuous(limits = c(3, 8), breaks = seq(0, 8, 1)) +
  scale_y_continuous(limits = c(0, 16), breaks = seq(0, 16, 2))
 
# plot both side by side
grid.arrange(alc1, alc2)
```

If we clean up our plot to only have average alcohol content, There is a small
increase once quality rating reaches 5. If the trend was better represented at
the lower ratings of 3 and 4, we could say this is a great correlation but as it 
is right now it may not be that significant. Let's explore some additional 
variables that also had higher values.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Volatile acidity and quality
ggplot(aes(quality, volatile.acidity), data = reds) +
  geom_point(alpha = 1/4, position = position_jitter(h = 0), color = 'blue') +
  geom_line(stat = 'summary', fun.y = mean, color = 'black') +
  xlab('Quality') +
  ylab('Volatile Acidity')
```

Volatile acidty and quality also had a solid correlation at -0.39. At lower 
values, ther's a nice line trending downwards showing that the more volatile
acidity in a wine, the lower rating it tends to received. Seems to flat line at
about a 7 quality rating.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# sulphates and wine quality
ggplot(aes(quality, sulphates), data = reds) +
  geom_point(alpha = 1/2, position = position_jitter(h = 0), 
             color = 'darkgreen') +
  geom_line(stat = 'summary', fun.y = mean, color = 'black') +
  xlab('Quality') +
  ylab('Sulphates')
```

Next I wanted to check out what other correlations looked like. The correlation
between sulphates and wine quality was one of the stronger ones for predictding
quality of wine at 0.25. The plot itself shows a modest incline of more sulphate
presence as wine quality increases. Sulphates are used to preserve wine and keep
the flavor fresh longer. Decent reflection of wine making but not a very
significant direct impact on wine quality rating.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# density and residual sugars
ggplot(aes(density, residual.sugar), data = reds) +
  geom_point(alpha = 1/4, position = position_jitter(h = 0), color = 'blue') +
  xlab('Density') +
  ylab('Residual Sugars')
cor.test(reds$density, reds$residual.sugar)
# doesn't seem to be a strong trend visually. density on x-axis probably not best
```

Density and residual sugar was another larger correlation at 0.35. Looking at
the scatterplot it's hard to tell the relation. I'm gonna try inverting the 
variables to see how that looks.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# same plot as above but inverting variables. 
ggplot(aes(residual.sugar, density), data = reds) +
  geom_point(alpha = 1/4, position = position_jitter(h = 0), color = 'blue') +
  scale_x_continuous(limits = c(0, 12), breaks = seq(0, 12, 2)) +
  scale_y_continuous() +
  xlab('Residual Sugar') +
  ylab('Density')
```

When you put residual sugar on the x-axis and density on the y-axis, we can 
start seeing a bit more of a trend but its still mostly a cluster. I thought 
this would be a visible relationship since more of a solid would probably mean
higher density. I also include similar limits to when we first looked at 
residual sugar to leave out some of the high value outliers.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# citric acid vs volatile acidity and fixed acidity
acid1 <- ggplot(aes(citric.acid, volatile.acidity), data = reds) +
  geom_point(alpha = 1/4, position = position_jitter(h = 0), color = 'yellow') +
  geom_line(stat = 'summary', fun.y = mean, color = 'black') +
  scale_x_continuous(limits = c(0, 0.75), breaks = seq(0, 0.75, 0.25)) +
  scale_y_continuous(limits = c(0, 1.25), breaks = seq(0, 1.25, 0.25)) +
  xlab('Citric Acid') +
  ylab('Volatile Acidity')
acid2 <- ggplot(aes(citric.acid, fixed.acidity), data = reds) +
  geom_point(alpha = 1/4, position = position_jitter(h = 0), color = 'yellow') +
  geom_line(stat = 'summary', fun.y = mean, color = 'black') +
  scale_x_continuous(limits = c(0, 0.75), breaks = seq(0, 0.75, 0.25)) +
  scale_y_continuous(limits = c(0, 14), breaks = seq(0, 14, 2)) +
  xlab('Citric Acid') +
  ylab('Fixed Acidity')

# plot both graphs side by side to contrast the trends
grid.arrange(acid1, acid2)
```

From our previous ggpairs function, I noticed some higher correlations between
citric acid and volatile acidty and fixed acidity. I'm unfamiliar with these variables so I did some research on the interaction between acetic acid and citric acid. 
Acetic acid is a volatile acid while citric is a fixed acid. More citric, more 
fixed acid which in previous plot showed a positive trend with quality. The 
opposite is true for acetic acid. Volatile acid has a negative correlation with 
quality.Here I plotted both relationships to see how fixed and volatile acidity 
interact with citric acid since we don't have values for acetic acid. Our 
findings from the data seem to support what I learned about citric acid and
acetic acid. More citric acid in a wine, higher fixed acidity values, r = 0.67. 
Less citric acid usually meant more volatile acidty but the correlation isn't as
strong, r = -0.55. So far these are some of the strongest correlation we've 
observed.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# ph and fixed acidity. These two variables should have a strong relationship. 
# more acid means lower pH value.
ggplot(aes(pH, fixed.acidity), data = reds) +
  geom_point(alpha = 1/4, position = position_jitter(h = 0), color = 'firebrick3') +
  scale_x_continuous(limits = c(3, 4.0), breaks = seq(3, 4.0, 0.2)) +
  xlab('Ph Level') +
  ylab('Fixed Acidity')
cor.test(reds$pH, reds$fixed.acidity)
```

I explored ph and fixed acidity since it was another strong correlation at -0.68
and its pretty straight forward. More acidity usually means lower pH values.
There is a nice downward trend between these two variables.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

Volatile acidity and quality was a really interesting one. One the largest 
negative correlations in our data set. After some quick research, I found out
that the primary volatile acid in wine is acetic acid which is associated with
the taste and smell of vinegar. This begins to explain the correlation. More
volatile acid means more smelly and pungent wine which could lower quality.

After reading more about sulphates(sulphites) in wine, sulphates are used 
primarily as a preservative to prevent oxidization and keep the wine fresh. 
This is a reflection of better brewed and treated wines which would indirectly
increase the quality of wine over various variables, not just sulphates.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

Density and residual sugar had a correlation of 0.35 which is the strongest
correlation that residual sugar has compared to other variables. Otherwise, 
residual sugar doesn't have other significant interactions. This makes sense
since red wine tends to have low sugar and isn't a key component to good wine.

Density also has strong correlations with fixed acidity, and citric acid. These
correlations makes sense since density is a measure of content per volume.
More sugar or acidic material would make it denser than water.

Citric acid has a positive correlation with fixed acidity and negative 
correlation with volatile acidity. I'm not an expert on acids so I did some 
reading. Citric acid added before the fermentation process can turn into
acetic acid which is the most common source for volatile acidity and explains
the correlation. There is less citric acid because it turns into acetic acid 
which adds to volatile acidity. If added after the fermentation process, citric
acid can complement a specific flavor and also add to its 'fresh' flavor.

### What was the strongest relationship you found?

The strongest correlation I found was the negative correlation between pH
and fixed acidity, r = -0.68. This one is straight forward since acid levels are 
on the lower end of the pH scale, 1-6, with neutral being at 7. Most of our wine
falls within the 3-4 which is a mild acid.

# Multivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots}
#adding a variable to data set to look at buckets of quality
reds$quality.bucket <- cut(reds$quality,
                             c(2, 4, 6, 8))

# alcohol and citric acid have sizeable correlations with quality
# alcohol and citric acid and quality
ggplot(aes(x = alcohol, y = citric.acid), data = subset(reds, quality>=3)) +
  geom_point(alpha = 0.6, position = 'jitter', size = 2, aes(color = quality.bucket)) +
  scale_color_manual(values=wes_palette(n=3, name="Zissou")) +
  xlab('Alcohol Content') +
  ylab('Citric Acid')
# higher quality wines in yellow cluster in top right of plot
```

As we explore interactions between mutliple variables, I grouped wine quality
into three buckets to make it easier to see the interaction with alcohol content
and citric acid. I picked these two variables since they had the two highest 
correlations with quality. In this plot, high quality wine is shown in yellow. 
You can begin to see this cluster hang in the top right quadrant suggesting that 
high quality wine tends to be higher in alcohol content and citric acid.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Second look at plot but this time with two buckets to see how it relfects the
# peaks found in initial quality plot

# Two buckets to separate 'high' quality wine and 'low' quality wine
reds$quality.bucket <- cut(reds$quality,
                             c(2, 5, 8))

# plotting again but two groups instead
ggplot(aes(x = alcohol, y = citric.acid), data = subset(reds, quality>=3)) +
  geom_point(alpha = 0.6, position = 'jitter', size = 2, aes(color = quality.bucket)) +
  scale_color_manual(values=wes_palette(n=3, name="Royal1")) +
  xlab('Alcohol Content') +
  ylab('Citric Acid')
# lower quality hangs in bottom left quadrant while high hangs out in top right
```

Since we had two peaks right around the middle, 5 and 6 rating, I made buckets
to reflect the two sides of the peak and see if the clusters can show any major
differences between the two ends of this data set. Here the data is a little bit
random but wine with a rating higher than 5 seems to be higher in wine quality
making up the majority of our points right of the 11% mark

```{r echo=FALSE, message=FALSE, warning=FALSE}
# density vs residual sugar vs fixed acidity
ggplot(reds, aes(x = fixed.acidity, y = residual.sugar, color= density)) +
  geom_point(alpha = 0.5, position = 'jitter') +
  xlab('Fixed Acidity') +
  ylab('Residual Sugar')
```

Both residual sugar and fixed acidity were correlated to higher density values.
When we plot all three variables, we can see the two correlations compliment
each other. Values high in fixed acidity and residual sugar tend to have a 
lighter shade of blue indicating higher density than the dark blue points Most 
of the darker blue points hang in the bottom left of our plot indicating lower
values in fixed acidity or residual sugar.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Building linear model to predict quality with 5 variables
r1 <- lm(I(quality) ~ I(alcohol), data = reds)
r2 <- update(r1, ~ . + citric.acid)
r3 <- update(r2, ~ . + volatile.acidity)
r4 <- update(r3, ~ . + sulphates)
r5 <- update(r4, ~ . + density)
mtable(r1, r2, r3, r4, r5, sdigits = 3)
# r-squared values are about 0.3. Not very strong compared to diamond set where
# we had high 0.9's
```

Using what I learned from the EDA class on linear regression, I made a linear 
regression model to try to predict quality with a few more variables. I chose 
alcohol, citric acid, volatile acidity, sulphates, and density because they had 
the highest correlations to quality. While the r squared values are fairly low, 
this model could help when picking out wine if you know a few of the variables 
for that particular wine to guess how good in quality it might be.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# attempt a prediction of quality using alchol, c-acid, v-acidity, sulphates, 
# and density
myWine = data.frame(alcohol = 13, citric.acid = 0.54, volatile.acidity = 0.31, 
                    sulphates = 0.82, density = 0.998)
modelEstimate = predict(r5, newdata = myWine, interval = "prediction", level = 0.95)
modelEstimate
```

To test out our linear regression model I provided a wine with hypothetical wine
descriptors, not real values, to fill out all 5 of the variables I used in my r5 
model. Our model predicted that the quality of our hypothetical wine would be 
about 6.84 with a lower bound of 5.54 and upper bound of 8.14. This seems 
reasonable since I tried to fill out some hypotheticalvalues that would be rated
on the higher end of their respective category. It's important to remember our 
low correlation values so I wouldn't put too much faith on these predictions.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# real world wine I got on a Friday night. Only had access to alcohol level
chateau = data.frame(alcohol = 14.3)
mod2 = predict(r1, newdata = chateau, interval = "prediction", level = 0.95)
mod2
# estimate seems reasonable and real world taste makes sense!
```

I did a real world test by grabbing a bottle of wine from a nearby shop trying
to get a bottle high in alcohol. I found a bottle with 14.3% alcohol content. 
Using my r1 model, just alcohol to predict quality, we get a predicted quality
rating of 7.04 which is fairly high, lower 5.64 and upper 8.43. This is 
fairly reasonable with our limited variable descriptor. The only way to get more
info was if I had a chemistry kit to examine sulphates level, acidity, etc. 
Tasted the wine and I would give it a 6 to 7 myself so I agree with the model in 
this example. I wouldn't expect the r1 model to be reliable in all future 
examples because we're only using one variable and the correlations are fairly 
low.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

The main takeaway from this project for me is learning how to identify good
wine with known variables. After plotting the quality scatter plot with
alcohol percentage on the x-axis and citric acid on the y-axis, I saw that if
you looked at more than one variable to predict quality, you start seeing some
nice patterns and are able to make some more consistent guesses.

### Were there any interesting or surprising interactions between features?

When you group qualities of wine into two groups, one low quality and other
high quality, you start to see clusters begin to separate when looking at
different variables but there is still a lot of overlap. I'm curious how 
price as a variable would interact with our observations in this data set. I
would guess it would behave similarly as our other variables in this data set
which is that it wouldn't have a strong impact on the quality. Generally more 
expensive would be better but it wouldn't be a 1 to 1 correlation.

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

I made a linear regression model using 5 variables to predict quality: alcohol,
citric acid, fixed acidity, sulphates, and density. I chose these variables 
because they had some of the stronger correlations to quality. Since all the r
squared value were generally low, I wouldn't put a lot of confidence in this 
model. Another drawback to this is that it isn't very practical to use 
on an everyday basis. You would need a chemistry kit that most wouldn't have
available in order to extract some of the variables needed to use this model
fully. Another limitation is the way that quality is rated on this data set. 
The highest rating given is an 8. With how this linear regression model works,
all you need is a couple outlier high values in a few of the variables to start
getting predicted quality values higher than 8 when we know that no wine was rated
higher than 8. Perhaps this is an issue with the critics in the data set but
as it stands the linear regression model doesn't do much to counter act
potentially giving a value higher than our max reported quality values.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}
# quality and alcohol. Plotting mean alcohol across quality includes Q1 and Q3
# also plotting geom smooth line to see gradual incrase in alcoholMean content
ggplot(aes(factor(quality), alcohol), data = reds) +
  geom_boxplot(color = 'firebrick3') +
  xlab('Quality, Score Between 0 and 10') +
  ylab('Alcohol Content (% by volume)') +
  ggtitle('Quality vs Alcohol Content')

```

### Description One
This plot was one of the most surprising to me. I normally had associated 
heavier alcohol content with poor taste but this plot showed a completely
opposite trend. It's a nice and clear plot showing that as alcohol percentage
increase, quality also tends to increase. While the trend isn't clear for values
lower than and including 4, there is a positive trend from 5 and upwards.

### Plot Two
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}
# citric acid vs volatile acidity and fixed acidity
acid1 <- ggplot(aes(citric.acid, volatile.acidity), data = reds) +
  geom_point(alpha = 1/4, position = position_jitter(h = 0), color = 'yellow') +
  geom_line(stat = 'summary', fun.y = mean, color = 'black') +
  ggtitle("Citric Acid's Correlation with Volatile and Fixed Acidity") +
  scale_x_continuous(limits = c(0, 0.75), breaks = seq(0, 0.75, 0.25)) +
  scale_y_continuous(limits = c(0, 1.25), breaks = seq(0, 1.25, 0.25)) +
  xlab('Citric Acid (g / dm^3)') +
  ylab('Volatile Acidity (g / dm^3)') 
  
acid2 <- ggplot(aes(citric.acid, fixed.acidity), data = reds) +
  geom_point(alpha = 1/4, position = position_jitter(h = 0), color = 'yellow') +
  geom_line(stat = 'summary', fun.y = mean, color = 'black') +
  scale_x_continuous(limits = c(0, 0.75), breaks = seq(0, 0.75, 0.25)) +
  scale_y_continuous(limits = c(0, 14), breaks = seq(0, 14, 2)) +
  xlab('Citric Acid (g / dm^3)') +
  ylab('Fixed Acidity (g / dm^3)')

# plot both graphs side by side to contrast the trends
grid.arrange(acid1, acid2)
```

### Description Two

This plot summarizes the interaction between citric acid and fixed/volatile 
acidity. Citric acid is primarily a fixed acid while acetic acid, it's counter-
part, is primarily a volatile acid. This is because citric acid turns into 
acetic acid during the fermentation process which would increase the volatile
acidity while lowering fixed acidity. There's an inverse relationship here. More
citric acid usually means lower acetic acid and lower volatile acidity.

### Plot Three
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}
#adding a variable to data set to look at buckets of quality
reds$quality.bucket <- cut(reds$quality,
                             c(2, 4, 6, 8))
# alcohol and citric acid along with quality
ggplot(aes(x = alcohol, y = citric.acid), data = subset(reds, quality>=3)) +
  geom_point(alpha = 0.6, position = 'jitter', size = 2, aes(color = quality.bucket)) +
  scale_color_manual(values=wes_palette(n=3, name="Zissou")) +
  xlab('Alcohol Content (% By Volume)') +
  ylab('Citric Acid (g / dm^3)') +
  ggtitle('Alcohol Content and Citric Acid correlation on Quality')

```

### Description Three

This last plot is a nice summary of my findings. It demonstrates the surprise
alcohol and quality correlation while also including the positive effect of 
citric acid. This plot also separates wine into three tiers, low, middle, and
high quality wine. I chose this color palette because it was really easy to see
the high quality wines compared to low and medium quality. There is a nice 
cluster of blues for low and mid quality wine with the yellow clusters showing
the high quality wine hanging out in the top right quadrant, indicating that
high quality wine generally is high in alcohol content and also high in citric 
acid compared to other wines.
------

# Reflection

After spending several hours exploring this red wine data set I feel that I've
learned a lot about wine using my newly acquired EDA skills. I was surprised by
the alcohol positive correlation on quality. I learned a bit about the 
fermentation process and how citric flavors could mean better tasting wine. Some
struggles included not knowing where to start. When you're unfamiliar with a 
data set, it's hard to start somewhere. The ggpairs function was super helpful
for this since I can get an overview and beging to dig. Another struggle was
playing with color palettes. Some times the colors were too similar and data
points were clustered too closely together. I spent a lot of time playing with
colors, alpha sizes, jitter, etc. It was also tougher than expected to figure 
out how to manipulate some of these variables to make plots that were easy to
extract trends from. Overall I feel like I learned a decent amount about wine
and consider myself more of a wine enthusiast. For future projects I would like
to strengthen my wrangling skils so I can work with variables a bit more in
depth. I had some trouble plotting variables with certain functions because I 
was accidentally putting quantitative variables into qualitative functions but
then the plots wouldn't come out like I wanted, etc. 